Elephant Seal Age JAE Intraspecific Variation

### 2023-11-27 paper/paperv2.qmd
Updating to post-1994

sealdat <- read_csv(here::here("data/raw/128L pull 2023_11_21.csv"), 
                    show_col_types = FALSE) %>% 
  mutate(observed = if_else(observed == "B", "Breeder", "Non-breeder"), 
         observed_int = if_else(observed == "Breeder", 1, 0), 
         pup_survived = pupseeneveragain > 0) %>% 
  filter(age > 3)

@@

breeding_mod <- glmer(observed_int ~ poly(age, 2) + (1 | animalID) + (1 | year),
                      data = filter(sealdat, year >= 1990), 
                      family = "binomial")
breed_pred <- ggpredict(breeding_mod, terms = c("age [all]"))
summary(breeding_mod)

@@

age_dist <- sealdat %>% 
  filter(age < 22,
         year >= 1990) %>% 
  group_by(age) %>% 
  count(age) %>% 
  ggplot(aes(x = age, y = n)) + 
  geom_col(fill = "#829BD1") + 
  labs(x = "Age (Years)", y = "Number of individuals") + 
  theme_few() + 
  theme(axis.title = element_text(size = 13))

@@

breed_nonbreed <- sealdat %>% 
  filter(year >= 1990) %>% 
  group_by(age) %>% 
  summarize(n_observed = sum(!is.na(observed)), 
            perc_breed = sum(observed == "Breeder") / n_observed, 
            se = (perc_breed * (1 - perc_breed) / n_observed)^0.5) %>% 
  filter(age < 22) %>%
  ggplot(aes(x = age, y = perc_breed)) + 
  geom_ribbon(aes(x, ymin = conf.low, ymax = conf.high),
              breed_pred,
              inherit.aes = FALSE,
              fill = "#829BD1",
              alpha = 0.25) +
  geom_line(aes(x, predicted),
            breed_pred,
            inherit.aes = FALSE,
            color = "#829BD1",
            linewidth = 1) +
  geom_point() + 
  geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), width = 0) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = "Maternal Age (Years)", y = "% Breeders (Mean ± SE)") + 
  theme_few() + 
  theme(axis.title = element_text(size = 13))

>>>>

sealdat <- read_csv(here::here("data/raw/128L pull 2023_11_21.csv"), 
                    show_col_types = FALSE) %>% 
  mutate(observed = if_else(observed == "B", "Breeder", "Non-breeder"), 
         observed_int = if_else(observed == "Breeder", 1, 0), 
         pup_survived = pupseeneveragain > 0) %>% 
  filter(age > 3, 
         year > 1994)

@@

breeding_mod <- glmer(observed_int ~ poly(age, 2) + (1 | animalID) + (1 | year),
                      data = filter(sealdat), 
                      family = "binomial")
breed_pred <- ggpredict(breeding_mod, terms = c("age [all]"))
summary(breeding_mod)

@@

age_dist <- sealdat %>% 
  filter(age < 22) %>% 
  group_by(age) %>% 
  count(age) %>% 
  ggplot(aes(x = age, y = n)) + 
  geom_col(fill = "#829BD1") + 
  labs(x = "Age (Years)", y = "Number of individuals") + 
  theme_few() + 
  theme(axis.title = element_text(size = 13))

@@

breed_nonbreed <- sealdat %>% 
  group_by(age) %>% 
  summarize(n_observed = sum(!is.na(observed)), 
            perc_breed = sum(observed == "Breeder") / n_observed, 
            se = (perc_breed * (1 - perc_breed) / n_observed)^0.5) %>% 
  filter(age < 22) %>%
  ggplot(aes(x = age, y = perc_breed)) + 
  geom_ribbon(aes(x, ymin = conf.low, ymax = conf.high),
              breed_pred,
              inherit.aes = FALSE,
              fill = "#829BD1",
              alpha = 0.25) +
  geom_line(aes(x, predicted),
            breed_pred,
            inherit.aes = FALSE,
            color = "#829BD1",
            linewidth = 1) +
  geom_point() + 
  geom_errorbar(aes(ymin = perc_breed - se, ymax = perc_breed + se), width = 0) +
  scale_y_continuous(labels = scales::percent, limits = c(0,1)) + 
  labs(x = "Maternal Age (Years)", y = "% Breeders (Mean ± SE)") + 
  theme_few() + 
  theme(axis.title = element_text(size = 13))
